Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-07 cluster_paper/ms.tex
update ms

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  It is
  important to verify that these ``halos'', ``tails'', and ``strings''
  are of similar ages and compositions as stars in the denser part of
  the cluster.  We present an analysis of NGC\,2516
  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
  and an apparent halo of stars spanning 500\,pc.  Combining
  photometry from Gaia, rotation periods from TESS, and lithium
  measurements from Gaia-ESO and GALAH, we find that the halo of
  NGC\,2516 is real and coeval with the core.  Two thirds of
  kinematically selected halo members out to 250\,pc from the cluster
  center have rotation periods consistent with a gyrochronological age
  of 150\,Myr.  A comparison sample of field stars shows no such
  trend.  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of
  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
  paradigm in which the halos of many open clusters are more populous
  than their cores.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

Our goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for the more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt a PARSEC-based isochrone age for \cn\ of $165\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of the statistical
and systematic components from \cite{cummings_2018}. 
%The relative age of ``slightly older than
%the Pleiades'' is likely the more robust statement, consistent with
%reports since \citet{mermilliod_comparative_1981}.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
    (see Section~\ref{subsubsec:cluster}).  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $\bpmrp$$<$1.2.
    From $1.2<\bpmrp<1.7$, 
    some NGC\,2516 members have longer rotation periods than Pleaides members.  
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
			\includegraphics[width=1\textwidth]{f5a.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with
  Gaia-ESO ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH
  spectra comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  The gray points are
  Keck-HIRES measurements of stars with detected transiting planets
  from the Kepler mission \citep{berger_identifying_2018}.  Points
  with ${\rm EW}\approx0\,$m\AA\ are non-detections.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
		\includegraphics[width=1\textwidth]{f5b.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Lithium and rotation in NGC\,2516.}
    Points show all kinematic \cn\ members with TESS rotation periods
    in the Gaia-ESO or GALAH surveys, and are colored by equivalent
    width.  Points with reported equivalent widths below 20\,m\AA\
    (including non-detections) are shown with gray crosses.  The
    lithium-rotation correlation for the K-dwarfs has been observed in
    other young clusters, and is discussed in
    Section~\ref{subsec:lithium}.
		\label{fig:lithiumrot}
	}
\end{figure*}

The third and final approach we took for confirming the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  The method has been reviewed by
\citet{soderblom_ages_2010}, and its connection to stellar rotation
has recently been surveyed by \citet{bouvier_lithium-rotation_2020}.
For \cn, two spectroscopic datasets seemed particularly important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant, respectively \citep[{\it
e.g.},][]{randich_gaiaeso_2018,buder_galah_2020}.  The target
selection and results from each survey were as follows.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<\bpmrp<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall increase of Li EW from
$0<\bpmrp<1.0$ is thought to be driven by the temperatures in the
stellar chromospheres: hotter stars fully ionize Li out of its ground
state \citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.},
later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
K dwarfs.  For comparison, we have also plotted the Li EWs measured by
\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
field (we interpolated from their effective temperatures using the
\citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2019.3.22.}.  The field star distribution peaks at late F-type
stars.  The majority of kinematically selected stars in the core and
the halo show lithium equivalent widths substantially in excess of
these field stars.

@@

\subsection{How did the halo form?}
\label{subsec:origin}
% Should we be calling the halo a tail?

Any theory for the structure of \cn, the Psc\,Eri stream, and
the strings and halos found by
\citetalias{kounkel_untangling_2019} and \citetalias{meingast_2021}
needs to explain a few observations.  First is their aspect
ratios.  The halo of \cn, for instance,
extends over at least 500\,pc, despite being only $\approx$25\,pc
wide.  Second is their axial tilts in the plane of the Milky Way -- why
their leading and trailing arms tend to conform to the Galaxy's
differential rotation ({\it e.g.}, Figure~\ref{fig:XYZ} or Figure~13
of \citetalias{meingast_2021}).  Third is the correlation between the
mean LSR-subtracted cluster velocity and elongation axes (blue arrows in
Figure~\ref{fig:XYZ}).  Such a correlation has also been noted in Coma
Ber by \citet{tang_comaber_2019}.  The other halos may show
similar trends.

One possible explanation is that these structures are tidal tails
\citep[{\it e.g.},][]{krumholz_star_2019}. The idea would be that
stars escape out of the Lagrange points imposed by the Milky Way's
potential, and then form leading and lagging arms due to differential
rotation in the Galaxy.
In this context, Psc\,Eri for instance might represent the completion
of a process that \cn\ is still undergoing.

@@

% Possible observational directions could include analyzing whether the
% Gaia kinematics contain any evidence of cluster expansion,
% as well as analyzing the mass functions of stars in different
% regions of the clusters to determine whether they contain

% CITE ALSO: WRIGHT+2016, WRIGHT AND MAMAJEK 2018 -- LOOKED AT SCO CEN, FOUND
% THAT THEY DIDN'T COME FROM A SINGLE CLUSTER BEING DISRUPTED, BUT WERE MORE
% LIKELY EXPLAINED THROUGH ``EXTENDED'' STAR FORMATION OVER A MUCH LARGER VOLUME.
% 
% Similarly Zari et al 2019 studied the entire Orion complex, and found that
% different regions of the complex (one toward the Belt, one toward 25\,Ori)
% began forming stars roughly simultaneously. No evidence was found of it
% propagating from these initial sites.

% cometary-like cloud,
% oriented toward the Galactic plane, with two distinct components:
% a denser and enhanced star-forming (bent) Head, and a
% lower density and star-formation quieter 75 pc long Tail.

% ``The stars in these halos could have escaped from the cores of the
% cluster, or they could be remnants of a larger and lower-density star
% formation complex that has since dispersed.''
% Another possibility is {\it triggered star formation}.

% One relevant connection may be to the ``Mamajek-2'' stellar group
% (see Jilinski+2009).

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} have recently demonstrated that at
ages between 100 and 200\,Myr, young clusters ({\cn, Blanco\,1,
the Pleiades, Psc\,Eri, M\,35, and M\,50) all show similar rotation
period distributions.  The FGK stars show a ``slow sequence'', a
population of ``rapid rotators'', and in some cases stars in the
``void'' between the two populations ({\it e.g.,}, Psc\,Eri and \cn\
show more stars in the void than the Pleiades).  The late M dwarfs
with detected rotation periods tend to rotate very rapidly
($<3$\,days), although these are beyond the magnitude limit of our
analysis.  This section briefly discusses the origins of these
features in the rotation-color diagram.

@@

A separate effect is that on the slow rotation sequence,
Figure~\ref{fig:rot}} shows that the binary stars appear to be either
preferentially redder, or to have faster rotation periods than the
single stars.  One likely explanation for this could be that the
unresolved binaries have a component contributing additional red light
to the system, skewing the color measurement of the primary.  Whether
any physical effects could be at play remains a question for future
work.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
been shown to carry over beyond just equivalent widths, and to the
actual abundance of lithium in the photosphere, A(Li).
\citet{soderblom_evolution_1993} calculated the relevant LTE curves of
growth for the Li\,6708\,\AA\ doublet, and used them to convert our
measured equivalent widths to abundances.  Non-LTE corrections affect
the abundances by $<0.5\,{\rm dex}$ and do not siginificantly affect
the global relation between Li abundance and stellar temperature or
rotation \citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation since
\citet{soderblom_evolution_1993} have expanded these results in the
Pleiades, the Psc-Eri stream, and M\,35, among other clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
In \cn\ specifically, \citet{jeffries_rotation_1998} observed a
similar trend that that seen in Figure~\ref{fig:lithiumrot} when
analyzing 24 stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
since these earlier studies.

@@

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?
{\bf FIXME: improve this paragraph}
%FIXME IMPROVE THIS PARAGRAPH.
Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

\subsection{Age of NGC\,2516}
\label{disc:absage}

Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
"""
Similar to Cummings et al. (2016), Cummings \& Kalirai (2018),
and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
\&
McCullough 2020)
"""

\paragraph{Gyrochronological Age}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrp<1.2$.  At redder colors
($1.2<\bpmrp<1.7$) the dispersion in rotation periods increases.
The maximum rotation periods seen in \cn\ across both sets
$\mathcal{A}$ and $\mathcal{B}$ extend up to $\approx$11 days, rather
than the $\approx$8.5day upper limit seen in the Pleiades.  This is
consistent with \cn\ being slightly older than the Pleiades.

Fitting a model to substantiate the claim that \cn\ is
gyrochronologically older than the Pleiades \citep[{\it e.g.}, any
of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
is, unfortunately, an exercise in tautology.  The aforementioned
models are empirically calibrated using the Pleiades and Praesepe.  In
other words, at $\bpmrp=1.35$, we observe rotation periods in \cn\
ranging from 7 to 10 days.  The 7 day rotation periods are consistent
with those observed in the Pleiades; the 10 day rotation periods are
not.  Fitting formulae from the previously cited studies give ages for
a star with $\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of 204
Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
\citet{spada_competing_2020} model has an absolute scale that is
well-calibrated at these ages and colors.  For an 8 day rotation
period of a star with the same color, this model quotes a 193 Myr age;
at 9 days, 388 Myr.  Given this degree of model uncertainty, we prefer
the relative statement that \cn\ appears slightly gyrochronologically
older than the Pleaides, and much younger than Praesepe.  Given a
Pleiades age of $130\pm15$\,Myr (an average of MSTO and
rotation-corrected lithium-depletion results;
\citealt{cummings_2018}), a gyrochronology-derived age estimate for \cn\
between 150 and 200\,Myr would therefore appear reasonable.  Much
older, and the issue of why the F and G dwarfs do not spin down
becomes pressing.

>>>>

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  It is
  important to verify that these ``halos'', ``tails'', and ``strings''
  are of similar ages and compositions as stars in the denser part of
  the cluster.  We present an analysis of NGC\,2516
  ($\approx$150\,Myr), which has a classical tidal radius of 10\,pc
  and an apparent halo of stars spanning 500\,pc.  Combining
  photometry from Gaia, rotation periods from TESS, and lithium
  measurements from Gaia-ESO and GALAH, we find that the halo of
  NGC\,2516 is real and coeval with the cluster's core.  Two thirds of
  kinematically selected halo members out to 250\,pc from the cluster
  center have rotation periods consistent with a gyrochronological age
  of 150\,Myr.  A comparison sample of field stars shows no such
  trend.  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of
  the Pleiades and Psc\,Eri.  Broadly speaking, this work advances the
  paradigm in which the halos of many open clusters are more populous
  than their cores.  We highlight the implications for spectroscopic
  survey targeting, open cluster dispersal, and planet searches around
  young stars.
\end{abstract}

@@

Our goal is to determine whether the ages of the core and halo
are consistent.  The left panel of Figure~\ref{fig:hr} suggests that
isochronally, they are: stars past the turnoff in both the halo and
core are on a consistent locus.  The right panel of
Figure~\ref{fig:hr} demonstrates the precision with which the claim
can be made.  The MSTO stars are consistent with the 178 and 316 Myr
models, but are bluer than the 100 Myr model.  The red-giant-branch (RGB) stars are
consistent only with the 178 Myr ($\log_{10} t/{\rm yr} = 8.25$)
model.  Based on the assumption that the width of the MSTO can be
attributed to binary stars, we are most interested in the blue edge
(blue stragglers are, however, a concern).  The blue edge appears most
compatible with the 178 Myr model.  These results are consistent with
the MSTO age range of 165 to 195\,Myr found by \citet{cummings_2018},
and we refer to their work for the more precise model-dependent
comparison.  Based on Table~3 of \citet{cummings_2018}, we therefore
adopt a MSTO age for \cn\ of $175\pm35$\,Myr, where
our quoted uncertainty is a quadrature sum of the statistical
and systematic components from \cite{cummings_2018}. 
%The relative age of ``slightly older than
%the Pleiades'' is likely the more robust statement, consistent with
%reports since \citet{mermilliod_comparative_1981}.

@@

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left}) and $\mathcal{B}$ ({\it right}) undergo successive stages of automated cleaning,
    (see Section~\ref{subsubsec:cluster}).  The Pleiades
    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 overlap with the Pleiades at $\bpmrp$$<$1.2.
    From $1.2<\bpmrp<1.7$, 
    some NGC\,2516 members have longer rotation periods than Pleaides members.  
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
			\includegraphics[width=1\textwidth]{f5a.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium in the core and halo of NGC\,2516.}
  Equivalent width of the 6708\,\AA\ doublet is shown versus
  dereddened color for all kinematic \cn\ members with Gaia-ESO
  ($N=452$) or GALAH ($N=107$) spectra available.  The GALAH spectra
  comprise slightly over half of the halo stars, due to the
  non-targeted selection function of that survey.  Field stars (gray
  points) are Keck-HIRES measurements of stars with detected
  transiting planets from the Kepler mission
  \citep{berger_identifying_2018}.  Stars in the Pleiades (125\,Myr)
  with Li detections are from \citet{bouvier_pleiades_lirot_2018};
  Praesepe detections (650\,Myr) are shown in violet
  \citep{soderblom_praesepe_1993}.  Points with ${\rm
  EW}\approx0\,$m\AA\ are non-detections.
  \label{fig:lithiumcorehalo}
  }
\end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		%\subfloat{
		\includegraphics[width=1\textwidth]{f5b.pdf}
		%	\includegraphics[width=0.59\textwidth]{f5b.pdf}
		%}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Lithium and rotation in NGC\,2516.} Points show all
  kinematic \cn\ members with TESS rotation periods in the Gaia-ESO or
  GALAH surveys, and are colored by equivalent width.  Points with
  reported equivalent widths below 20\,m\AA\ (including
  non-detections) are shown with gray crosses.  The lithium-rotation
  correlation for the K-dwarfs has been observed in other young
  clusters, and is discussed in Section~\ref{subsec:lithium}.
	\label{fig:lithiumrot}
	}
\end{figure*}

The third and final approach we took for confirming the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  This type of age measurement has been reviewed by
\citet{soderblom_ages_2010}, and its connection to stellar rotation
has recently been surveyed by \citet{bouvier_lithium-rotation_2020}.
For \cn, two spectroscopic datasets seemed particularly important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant, respectively \citep[{\it
e.g.},][]{randich_gaiaeso_2018,buder_galah_2020}.  The target
selection and results from each survey were as follows.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<\bpmrp<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall increase of Li EW from
$0<\bpmrp<1.0$ is driven by the temperatures in the stellar
chromospheres: hotter stars fully ionize Li out of its ground state
\citep[{\it e.g.}, Figure~4 of][]{soderblom_evolution_1993}.  The
depletion of stars redder than $\bpmrp\gtrsim1.4$, {\it i.e.}, later
than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier K
dwarfs: these stars have burned their lithium.  For comparison, we
have also plotted the Li EWs measured by
\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
field, and Pleiades members from \citet{bouvier_pleiades_lirot_2018}
(we interpolated from their effective temperatures using the
\citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2021.03.02.}.  For clarity, we have omitted the few upper
limits reported by \citet{bouvier_pleiades_lirot_2018}.  The field
star distribution peaks at late F-type stars.  The majority of
kinematically selected stars in the core and the halo show lithium
equivalent widths substantially in excess of these field stars.

@@

\subsection{How did the halo form?}
\label{subsec:origin}

Any theory for the structure of \cn, the Psc\,Eri stream, and the
strings and halos found by \citetalias{kounkel_untangling_2019} and
\citetalias{meingast_2021} needs to explain a few observations.  First
is their aspect ratios.  The halo of \cn, for instance, extends over
at least 500\,pc, despite being only $\approx$25\,pc wide.  Second is
their axial tilts in the plane of the Milky Way -- why their leading
and trailing arms tend to conform to the Galaxy's differential
rotation ({\it e.g.}, Figure~\ref{fig:XYZ} or Figure~13 of
\citetalias{meingast_2021}).  Third is the correlation between the
mean LSR-subtracted cluster velocity and elongation axes (blue arrows
in Figure~\ref{fig:XYZ}).  Such a correlation has also been noted in
Coma Ber by \citet{tang_comaber_2019}; the other halos may show
similar trends.

One possible explanation is that the observed halo structures are
tidal tails \citep[{\it e.g.},][]{krumholz_star_2019}. The idea would
be that as the cluster contracts, stars escape out of the Lagrange
points imposed by the Milky Way's potential, and form leading and
lagging arms due to differential rotation in the Galaxy.  Kinematic
evidence that the core of \cn\ may be undergoing contraction has
recently been presented by \citet{healy_stellar_2020}, and might
support this interpretation.  Whether the exact rate of contraction,
and the corresponding ``evaporation rate'' of the core is sufficient
to explain the formation of the halo has yet to be assessed.

@@

@@

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} have recently demonstrated that
between 100 and 200\,Myr, young clusters (\cn, Blanco\,1, Psc\,Eri, M\,35, M\,50, and the Pleaides)
all show similar rotation period
distributions.  The FGK stars show a ``slow sequence'', a population
of ``rapid rotators'', and in some cases stars in the ``void'' between
the two populations ({\it e.g.,}, Psc\,Eri and \cn\ show more stars in
the void than the Pleiades).  The late M dwarfs with detected rotation
periods tend to rotate very rapidly ($<3$\,days), although these are
beyond the magnitude limit of our analysis.  This section briefly
discusses the origins of these features in the rotation-color diagram.

@@

A separate effect is that on the slow rotation sequence,
Figure~\ref{fig:rot} shows that the binary stars appear to be either
preferentially redder, or to have faster rotation periods than the
single stars.  One likely explanation for this could be that the
unresolved binaries have a component contributing additional red light
to the system, skewing the color measurement of the primary.  Whether
any physical effects could be at play remains a question for future
work.

@@

The lithium-rotation correlation (Figure~\ref{fig:lithiumrot}) has
carries over beyond just equivalent widths, and to the actual
abundance of lithium in the photosphere, A(Li).  To our knowledge, the
first demonstration of this was by \citet{soderblom_evolution_1993},
who calculated the relevant LTE curves of growth for the
Li\,6708\,\AA\ doublet, and used them to convert equivalent widths to
abundances.  Non-LTE corrections affect the abundances by $<0.5\,{\rm
dex}$ and do not change the overall relationships between Li abundance
and stellar temperature or rotation
\citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation since
\citet{soderblom_evolution_1993} have expanded these results in the
Pleiades, the Psc-Eri stream, and M\,35, among other clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
In \cn, \citet{jeffries_rotation_1998} previously observed a
similar trend that that seen in Figure~\ref{fig:lithiumrot} when
analyzing 24 stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
since these earlier studies.

@@

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?
{\bf FIXME: improve this paragraph... (e.g., see the introductory
material in Jeffries+21}
%FIXME IMPROVE THIS PARAGRAPH.
Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

\subsection{The Age of NGC\,2516}
\label{disc:absage}

\paragraph{Age from Gyrochronology}
Comparing the slow sequence of the Pleiades and \cn\ more closely, the
top row of Figure~\ref{fig:rot} shows that the two sequences overlap
from $0.5<\bpmrp<1.2$.  At redder colors ($1.2<\bpmrp<1.7$) the
dispersion in rotation periods increases.  The maximum rotation
periods seen in \cn\ across both sets $\mathcal{A}$ and $\mathcal{B}$
extend up to $\approx$11 days, rather than the $\approx$8.5day upper
limit seen in the Pleiades.  This is consistent with \cn\ being
slightly older than the Pleiades.

Fitting a model to substantiate the claim that \cn\ is
gyrochronologically older than the Pleiades \citep[{\it e.g.}, any
of][]{mamajek_improved_2008,angus_toward_2019,spada_competing_2020}
is, unfortunately, an exercise in tautology.  Gyrochronological models
are empirically calibrated against the Pleiades and Praesepe.  In
other words, at $\bpmrp=1.35$, we observe rotation periods in \cn\
ranging from 7 to 10 days.  The 7 day rotation periods are consistent
with the Pleiades; the 10 day rotation periods are not.  Fitting
formulae from the previously cited studies give ages for a star with
$\{\bpmrp=1.35, (\mathrm{B}-\mathrm{V})_0=1.10, P=7\,{\rm day}\}$ of
204 Myr, 316 Myr, and 107 Myr respectively.  Therefore, only the
\citet{spada_competing_2020} model has an absolute scale that is
well-calibrated at these ages and colors.  For an 8 day rotation
period of a star with the same color, this model quotes a 193 Myr age;
at 9 days, 388 Myr.  Given this degree of model uncertainty, we prefer
the relative statement that \cn\ appears slightly gyrochronologically
older than the Pleaides, and much younger than Praesepe.  Given a
Pleiades age of $125\pm20$\,Myr (an average of MSTO and
rotation-corrected lithium-depletion results; see {\it e.g.}
\citealt{stauffer_keck_1998}, \citealt{soderblom_ages_2014}, and
\citealt{cummings_2018}), a gyrochronological age estimate for
\cn\ between 150 and 200\,Myr would therefore appear reasonable.  Much
older, and the issue of why the F and G dwarfs do not spin down
becomes pressing.

\paragraph{Age from Lithium Depletion}

The lithium depletion boundary for \cn\ has to our knowledge not been
measured.  Given the distance to the cluster, this is not surprising:
at $\sim$150\,Myr, the LDB is expected to occur at a stellar mass of
$\approx$0.088\,$M_\odot$ \citep{soderblom_ages_2014}.  This would
correspond to $\bpmrp\approx4.7$, well beyond the limits of available
cluster membership lists and spectroscopy.

An easier {\it relative} measurement to make of lithium depletion is
from the EW-color diagram (Figure~\ref{fig:lithiumcorehalo}).  The
\cn\ and Pleiades sequences in this diagram appear indistinguishable.
A relative lithium-based age estimate for \cn\ would therefore be
``Pleiades-age''.  The uncertainties on this estimate are large
because lithium depletion timescales are long for Sun-like stars past
100\,Myr.  For instance, \citet{jeffries_m35_li_2020} has shown in a
comparison of the Pleiades and M\,35 ($\approx$150\,Myr) that the two
are indistinguishable given current measurement, despite the age of
the latter being expected to be marginally older than the former.
\cn\ members are on average more depleted than {\it e.g.}, stars in
IC\,2602, but only marginally so \citep{soderblom_ages_2014}.  On the
older end though, Figure~\ref{fig:lithiumcorehalo} does show that
relative to Praesepe ($\approx$650\,Myr), the lithium EWs of \cn\ are
significantly enhanced.  Given these comparisons, a plausible range of
ages based on lithium for \cn\ would therefore be between $\approx$100
and 200\,Myr.

\paragraph{Adopted Age}

As noted in Section~\ref{subsec:hr}, a color-color analysis of the main
sequence turn-off (MSTO) by \citet{cummings_2018} found that
\cn\ is 40-60\,Myr older than the Pleiades.  Given a Pleiades age of
$125\pm20$\,Myr, this yields a MSTO age for \cn\ of
$175\pm35$\,Myr.  The gyrochronological age we have argued is likely
super-Pleiades, {\it i.e.}, in the range of 150 to 200\,Myr.  The
lithium age based on the depletion of the G and K dwarfs is more
uncertain, but is consistent with the Pleiades, and could be a bit
older ($\approx$100 to 200\,Myr).  Averaging these three different
indicators, the age of \cn\ would be $167\pm20$\,Myr, if
we set the uncertainties to match the absolute Pleiades age
uncertainty of $\pm 20\,$Myr \citep{soderblom_ages_2014}.  Whether
this absolute age for \cn\ is particularly meaningful, we leave the
reader to judge.
